Evolutionary rescue and reintroduction of resistant frogs allows recovery in the presence of a lethal fungal disease

### 2023-12-03 doc/manuscript/translocation.qmd
Describe stan_glmer models

### Population viability modeling

##### Incorporating yearly variability in vital rates

@@

##### Estimating model parameters

@@

##### Model analysis and simulation

@@

Second, we performed a local elasticity analysis on $\lambda$ with respect to parameters \$\\sigma\_{J_1}\$, $\sigma_{J_2}$, $\sigma_{A_R}$, and $F$ to determine how small changes in these parameters could influence the long-run deterministic growth rate of populations (@fig-viability-supp).

@@

>>>>

##### Methods

###### Incorporating yearly variability in vital rates

@@

###### Estimating model parameters

@@

###### Model analysis and simulation

@@

Second, we performed a local elasticity analysis on $\lambda$ with respect to parameters $\sigma_{J_1}$, $\sigma_{J_2}$, $\sigma_{A_R}$, and $F$ to determine how small changes in these parameters could influence the long-run deterministic growth rate of populations (@fig-viability-supp).

@@

#### Among-site survival modeling

We used Bayesian generalized linear mixed models to investigate predictors of survival among sites. 
The response $y_i$ is binary, representing a point estimate of whether individual $i$ survived in the year following translocation. 
We generated these point estimates by rounding the posterior median of 1-year post introduction survival for each individual, and modeled the data using a Bernoulli distribution:

$$y_i \sim \text{Bernoulli}(p_i),$$

where $p_i$ is the probability of survival.

We modeled variation in probabilities as follows:

$$\text{logit}(p_i) = \alpha + X_i \pmb \beta + \pmb \nu_{g[1:N]},$$

where $\alpha$ is an intercept, $X_i$ is a length $K$ row vector of predictors, $\pmb \beta$ is a column vector of predictor effects, and $\pmb \nu$ a vector of group level random effects. 
Here $g[i]$ refers to the group $g$ containing individual $i$, and we estimate an adjustment for each of the $G$ groups ($\nu_1, ..., \nu_G$).

These models were fit using the `stan_glmer()` function in the rstanarm package, with default priors described below [@rstanarm2022].
These priors are vague, but include data dependent scaling as follows to account for different input variable scales.
However, as we standardized all predictor variables similarly to have equal variance (by centering and dividing by twice the sample standard deviation), the resulting priors are identical. 
Specifically, we have:

$$\alpha \sim \text{Normal}(0, 2.5),$$
and

$$\beta_k \sim \text{Normal}(0, 5),$$

for $k=1, ..., K$ where $K$ is the number of predictor variables.

The default prior for group level adjustments $\nu_1, ..., \nu_G$ in rstanarm is a zero-mean Gaussian, where the covariance matrix is constructed from a correlation matrix with an LKJ prior, and a vector of variance parameters -- the decomposition of variance prior with unit regularization, concentration, shape, and scale parameters [@lewandowski2009].

We drew posterior samples using Dynamic Hamiltonian Monte Carlo in Stan, with four parallel chains, each run for 10,000 iterations, discarding the first half of each chain as warm-up draws [@rstanarm2022]. 
We used Rhat statistics and trace plots to verify convergence. 
We considered models with different subsets of fixed and random effects, and used approximate leave-one-out cross validation to identify the best model [@vehtari2016].
